Community-Conditioned Language Models

### 2020-11-18 paper/paper.tex
draft conclusion

In this work, we investigate linguistic variation across online communities.
Broadly construed, we want to investigate two questions.  First, we
are interested in variation in the level of linguistic complexity
exhibited by different communities.

TODO: insert some anecdotes (perhaps some particularly stupid/clever
comments from reddit here.)

It is folklore that certain communities are more linguistically complex than others.
Here we want to quantify this phenomenon.
More precisely, we investigate if the degree of complexity can be
captured in a computational model.  For this purpose define various
Community Conditioned Language Models (CCLM for short), which can be
modeled as an estimator $M$ of the degree to which a given message $m$ 
is acceptable in a given community $c$: \(P_M(m \mid c)\).

Second, we want to find out how, and how much, CCLMs can be attuned to
the specificity of communities. We test whether 1. the linguistic
models are correlated with a model of community based only on
co-occurence of users between communities and 2. if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language become and the easier it is to identify.

To sum up, we propose to test the following hypotheses:
\begin{hypotheses}
\item Different communities have different levels of linguistic
  complexity
\item The layer at which the community embedding is taken into account
  in CCLM influences its perplexity.
\item CCLM can be used to recognize communities based on
  latent bayesian classification.
\item The layer at which the community embedding is taken into account
  influences CCLM and its associated latent classifier.
  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
\item The model representation of communities are correlated with
  co-ocurrence of users between communities.
% \item Community stability is correlated with language complexity
\item Linguistically poor communities are easier to identify than linguistically rich communities.
\end{hypotheses}

\jp{TODO: Outline of the paper.}

@@

To investigate variation across communities, we use comments collected
from the social media website Reddit.\footnote{Comments were obtained
  from the archive at \url{https://pushshift.io/}.
  \cite{Baumgartner2020}.}

Reddit is divided into forums called \textit{subreddits}, 
which are typically organized around a topic of interest. 
Users create \textit{posts}, which consist of a link, image, 
or text, along with a \emph{comment} section. 
Comments are threaded: a comment can be made directly on a post,
or appear as a reply to another comment.
%
Hereafter we refer to such comments as ``messages'', to match our
convention in mathematical formulas: the letter $c$ stands for a
community, and $m$ stands for a message.

@@

We experiment with two kinds of model architecture: LSTM \citep{Hochreiter1997}
and Transformer \citep{Vaswani2017}.  In either case, the model
is organised as a standard $n$-layer neural sequence encoder with
token embedding (we set $n=3$ for all models). As is standard, words are predicted
using a softmax projection layer over a pre-determined vocabulary.
These encoders form the core of the community-conditioned language models (CCLM).

@@

We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
The vocabulary consists of the most frequent tokens at the message level (across all communities),
meaning that we only count each token once per message.
The main reason for this is to avoid over-counting otherwise rare tokens that were spammed
many times in a single message.

We trained the models on a simple left-to-right language modeling task
with cross entropy loss over the vocabulary. 
Since the transformer operates on all tokens in the sequence at once,
this was achieved by masking and incrementally un-masking input tokens
in a left-to-right fashion. \bn{does this address it?} %\jp{I don't think that this applies for the transformer model} 
We used the AdamW \citep{Loshchilov2019} optimization algorithm, 
with an initial learning rate of \num{0.001}.
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

@@

As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths .
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(Table \ref{tab:model-ppls}).

@@

\jp{Go through hypotheses again}

>>>>

In the present work, we investigate linguistic variation across online communities.
Broadly construed, we want to investigate two questions.  First, we
are interested in variation in the level of linguistic complexity
exhibited by different communities.

\jp{
  Some examples (perhaps include if there is space):

  (CFB)
  It was against SMU in the 47-48 Cotton bowl. They had cancelled a
  game the year before against Miami for the same reason. There's
  debate on if that's really the origin for the phrase but the story
  itself is true.

  (Videos)
  Reminds me of Y2K computer bug. A lot of work went into
  pre-emptively fixing it, and a lot of bugs were found and fixed,
  some of which would have been catastrophic. Unfortunately, because
  the work was successful, a lot of people now describe it as a waste
  of time/money and a big deal made out of nothing.  }

It is folklore that certain communities exhibit more linguistic complexity than others.
Here we want to quantify this phenomenon.
More precisely, we investigate if the degree of complexity can be
captured in a computational model.  For this purpose define various
Community Conditioned Language Models (CCLM for short), which can be
modeled as an estimator $M$ of the degree to which a given message $m$ 
is acceptable in a given community $c$: \(P_M(m \mid c)\).

Second, we want to find out how, and how much, CCLMs can be attuned to
the specificity of communities. We test whether 1. the linguistic
models are correlated with a model of communities based only on
co-occurence of users between communities and 2. if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language become and the easier it is to identify.

To sum up, we propose to test the following hypotheses:
\begin{hypotheses}
\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
  complexity
\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
  in a CCLM influences its perplexity.
\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
  latent bayesian classification.
% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
%   influences CCLM and its associated latent classifier.
  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
  co-ocurrence of users between communities.
% \item Community stability is correlated with language complexity
\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
  linguistically rich communities.
\end{hypotheses}

\jp{Perhaps: Outline of the paper.}

@@

We investigate linguistic variation across various communities present
on the social media website Reddit.\footnote{Comments were obtained
  from the archive at \url{https://pushshift.io/}.
  \cite{Baumgartner2020}.}
%
Reddit is divided into forums called \textit{subreddits}, 
which are typically organized around a topic of interest. 
Users create \textit{posts}, which consist of a link, image, 
or text, along with a \emph{comment} section. 
Comments are threaded: a comment can be made directly on a post,
or appear as a reply to another comment.
%
Hereafter we refer to such comments as ``messages'', to match our
convention in mathematical formulas: the letter $c$ stands for a
community, and $m$ stands for a message.

@@

We experiment with two kinds of model architecture: LSTM \citep{Hochreiter1997}\jp{Is it really a simple or a bi-lstm? Even If it's simple it may be worth specifying due to the prevalence of bilstm everywhere.}
and Transformer \citep{Vaswani2017}.  In either case, the model
is organised as a standard $n$-layer neural sequence encoder with
token embedding (we set $n=3$ for all models). As is standard, words are predicted
using a softmax projection layer over a pre-determined vocabulary.
These encoders form the core of the community-conditioned language models (CCLM).

@@

We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
The vocabulary consists of the most frequent tokens at the message level (across all communities),
meaning that we only count each token once per message.
The main reason for this choice is to avoid over-counting otherwise rare tokens that were spammed
many times in a single message.

For the LSTM, we trained the models on a simple left-to-right language modeling task
with cross entropy loss over the vocabulary. 
Because the transformer operates on all tokens in the sequence at once,
the language modeling was achieved by masking and incrementally un-masking input tokens
in a left-to-right fashion, following \citet{BERT}\jp{Or some other standard paper}.  
We used the AdamW \citep{Loshchilov2019} optimization algorithm, 
with an initial learning rate of \num{0.001}.\jp{Also for the transformer? No simulated annealing?}
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

@@

As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths.
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(Table \ref{tab:model-ppls}).

@@

We find that various communities exhibit widely varying levels of
linguistic complexity (\cref{hyp:varying-complexity}), with perplexity
per-word ranging from less than 53 to 104 for the best model
architecture.

With indiscernibility values of 1.5 to 3.5 (out of a maximum of 46),
we find that CCLMs can be used to recognize communities based on
latent bayesian classification (\cref{hyp:LMCC-works}). Even the most
generic communities appear to exhibit fairly specific linguistic
patterns. Furthermore, these patterns are just as recognizable by LSTM
or Transformer models, when properly configured and trained.

We find a weak correlation (TODO; r , p) between the perplexity of the
CCLM and the indiscernibility of the corresponding community
(\cref{hyp:rich-harder-to-identify}). Because we have relatively few,
and only large communities in our test set, it is difficult to
provide a conclusive answer, but this kind of correlation would suggest that
community-specific complexity can be partially attributed to using
more varied patterns. However, even the most complex communities
(perplexity per word above 100) have an indiscernibility of about 3
(out of a maximum of 46), so they retain fairly specific patterns:
they are not simply an even mixture of the kind of comments found in
other communities.

We find that the layer at which the community embedding is taken into
account in a CCLM influences its perplexity (\cref{hyp:layer-effect}),
albeit weakly. For LSTM models, the perplexity per word, averaged over
all communities, varies between 78.64 and 80.70. It is steadily
increasing when $l_c$ decreases.  For transformer models, it varies
between 89.38 and 122.56. Here perplexity per-word is not a monotonous
function of $l_c$: the maximum is reached at $l_c=2$. We did not
expect non-mononicity, and leave studying this effect to future work.

The representation of communities built by CCLM is exhibiting a clear
correlation with user-coocurrence patterns
(\cref{hyp:extra-linguistic-correlation}, TODO; r , p). A natural
question that may arise is how much this correlation can be explain by
user-specifc linguistic patterns. We plan to control for this variable
in future work, for example by building author-conditioned language
models and testing if community vector embeddings can be determined by
the set of the vector embeddings of its members.
